A large and variable leading tail of Helium in a sub-Saturn undergoing runaway inflation

### 2023-06-05 paper/paper1/ms.tex
Comment-out the How We Constructed the Contours discussion, which was probably too wordy and pedantic

Under either scenario, HAT-P-67b resides in an extremely short-lifetime region.  HAT-P-32b offers an interesting test case since it has recently been shown to exhibit a large helium excess \authorcomment1{cite ZJ}.  It resides just slightly denser than the 0.1 $g\;cm^{-3}$ dividing line, with XUV predicting over 10$\times$ longer inflationary timescale for HAT-P-32b than for HAT-P-67b, while Ohmic dissipation expects merely a factor of 3 difference.

One other key observation may tip the scales in favor of XUV irradiation.  The conspicuous leading tail favors preferential mass loss on the highly irradiated planet dayside.  It is not clear to what extent Ohmic dissipation can imbue preferential dayside mass loss, and on balance, we attribute the observation of a large leading tail as supporting the XUV-irradiation-driven mass loss scenario.

Finally, we consider the prospect that the planetary mass is much lower than the 1-$\sigma$ estimate, such that the white-light planetary radius \emph{is} the Hill radius (we have previously assumed the Hill radius is 2.7 planetary radii).  This extreme scenario could manifest enormous mass loss rates, with the atmosphere's reservoir of mass fleeing the gravitational potential well directly, without the cushion of an exosphere.  This terminal Roche Lobe overflow should have observational consequences.  In particular, heavy elements would easily leak into the planetary wind, yielding possibly many observable metal lines in the UV.

@@

As seen in Figure \ref{fig:ThornFortUpdated}, \hatpb~has few other known analogs.  Figure \ref{fig:tf2018_theory} illustrates why: the inflationary lifetime for planets with comparable mass, radius, and $T_\mathrm{eq}$ is exceptionally short, and so they are rare.  We constructed the $\tau_\mathrm{infl}$ contours in the following way.  We censored the $\tau_\mathrm{infl}$ to regions in the mass-radius plane where numerical simulations of Ohmic dissipation resulted in runaway inflation \citet{2011ApJ...738....1B}.  We interpolate the $T_\mathrm{eq}$ trendlines from Fig. \ref{fig:ThornFortUpdated} to compute $\tau_\mathrm{infl}$ from Equation \ref{eqInflate}, including the best fit Gaussian prescription for $\hat{\epsilon}_G$ from Eq. 34 of \citet{2018AJ....155..214T}.  This trend should exhibit significant uncertainty owing to the order-of-magnitude nature of the scaling relation, source-to-source scatter in $\hat{\epsilon}_G$, and to what extent $\hat{\epsilon}_G$ applies for planets less than $0.5 M_\mathrm{Jup}$, which were excluded in the derivation of $\hat{\epsilon}_G$ owing to their scarcity.  Nevertheless, this contour appears to match the bulk trends: planets tend to reside in either the non-inflationary, stable locus or the inflationary-yet-stable locus.  A relatively few $\sim$dozens of planets approach the ``cosmic shoreline'' of inflationary lifetimes in the tens of Myr. The detection of atmospheric escape along this frontier (green circles in Figure \ref{fig:tf2018_theory}) supports this interpretation.  More observations and theoretical modeling of atmospheric escape along this ridge could help better understand its realized versus expected shape.

>>>>

Under either scenario, HAT-P-67b resides in an extremely short-lifetime region.  HAT-P-32b offers an interesting test case since it has recently been shown to exhibit a large helium excess \authorcomment1{cite ZJ}.  It resides just slightly denser than the 0.1 $g\;cm^{-3}$ dividing line, with XUV predicting over 10$\times$ longer inflationary timescale for HAT-P-32b than for HAT-P-67b, while Ohmic dissipation expects merely a factor of 3 difference.  HAT-P-32b exhibits a more symmetric leading and trailing tail, whereas HAT-P-67b stands out as exhibiting evidence for preferential mass loss on the highly irradiated planet dayside.  These differences may make this pair an especially valuable laboratory for discriminating among theories of atmospheric escape.

Finally, we consider the prospect that the planetary mass of HAT-P67b is much lower than the 1-$\sigma$ estimate, such that the white-light planetary radius \emph{is} the Hill radius (we have previously assumed the Hill radius is 2.7 planetary radii).  This extreme scenario could manifest enormous mass loss rates, with the atmosphere's reservoir of mass fleeing the gravitational potential well directly, without the cushion of an exosphere.  This terminal Roche Lobe overflow should have observational consequences.  In particular, heavy elements would easily leak into the planetary wind, yielding possibly many observable metal lines in the UV.

@@

%We constructed the $\tau_\mathrm{infl}$ contours in the following way.  We censored the $\tau_\mathrm{infl}$ to regions in the mass-radius plane where numerical simulations of Ohmic dissipation resulted in runaway inflation \citet{2011ApJ...738....1B}.  We interpolate the $T_\mathrm{eq}$ trendlines from Fig. \ref{fig:ThornFortUpdated} to compute $\tau_\mathrm{infl}$ from Equation \ref{eqInflate}, including the best fit Gaussian prescription for $\hat{\epsilon}_G$ from Eq. 34 of \citet{2018AJ....155..214T}.  This trend should exhibit significant uncertainty owing to the order-of-magnitude nature of the scaling relation, source-to-source scatter in $\hat{\epsilon}_G$, and to what extent $\hat{\epsilon}_G$ applies for planets less than $0.5 M_\mathrm{Jup}$, which were excluded in the derivation of $\hat{\epsilon}_G$ owing to their scarcity.  Nevertheless, this contour appears to match the bulk trends: planets tend to reside in either the non-inflationary, stable locus or the inflationary-yet-stable locus.  A relatively few $\sim$dozens of planets approach the ``cosmic shoreline'' of inflationary lifetimes in the tens of Myr. The detection of atmospheric escape along this frontier (green circles in Figure \ref{fig:tf2018_theory}) supports this interpretation.  More observations and theoretical modeling of atmospheric escape along this ridge could help better understand its realized versus expected shape.
